### 2017-07-10 chapters/paper1/article.tex
Update to latest version

% \documentclass[a4paper,10pt]{article}
% % \documentclass[times]{cnmauth}
% \usepackage[utf8]{inputenc}
% \usepackage{amsmath}
% \usepackage{amssymb}
% \usepackage{caption}
% \usepackage{subcaption}
% \usepackage{graphicx}
% \usepackage{tikz}

% \usepackage[noend]{algpseudocode}
% \usepackage{algorithm}
% \usepackage{enumitem}% http://ctan.org/pkg/enumitem
% \usepackage{color}
% \usepackage{url}
% \usepackage{booktabs}
% \usepackage[export]{adjustbox}
% % \usepackage[tablesonly]{endfloat} %Data table put at end according to Wiley preference.
\graphicspath{{chapters/paper1/figures/}}

@@

% \usetikzlibrary{shapes,arrows}
% \usetikzlibrary{positioning}

@@

% \def\baselinestretch{2}

% \begin{document}

%opening
% \title{Dyssnychrony Specific Personalization of a Cardiac Mechanics Model by Adjoint Optimization}
% \title{High Resolution Data Assimilation of Cardiac Mechanics \\ \large Applied to a Dyssynchronous Ventricle}
 % \author{Gabriel Balaban\footnotemark \affil{1}~\affil{2}~\affil{5},
 %         Henrik Finsberg\footnotemark[\value{footnote}]~\affil{1}~\affil{2}~\affil{5},
 %         Hans Henrik Odland~\affil{4}~\affil{7}, 
 %         Marie~E.~Rognes~\affil{1}~\affil{3},
 %         Stian Ross~\affil{4}~\affil{5},
 %         Joakim Sundnes~\affil{1}~\affil{2}~\affil{5},
 %         Samuel Wall~\affil{1}~\affil{6}~\affil{5}\\}

% \address{\affilnum{1} Simula Research Laboratory, P.O. Box 134 1325 Lysaker, Norway \break
% \affilnum{2}Department of Informatics, University of Oslo, P.O. Box 1080, Blindern 0316 Oslo, Norway \break
% \affilnum{3}Department of Mathematics, University of Oslo, P.O Box 1053, Blindern 0316 Oslo, Norway \break
% \affilnum{4}Faculty of Medicine, University of Oslo, P.O. Box 1078 Blindern, 0316 Oslo, Norway \break
% \affilnum{5}Center for Cardiological Innovation, Songsvannsveien 9, 0372 Oslo, Norway \break
% \affilnum{6}Department of Mathematical Science and Technology, Norwegian University of Life Sciences,
% 	 Universitetstunet 3 1430 Ås, Norway \break
% \affilnum{7}Department of Pediatrics, Oslo University Hospital, PO Nydalen, Oslo, Norway}

% \corraddr{gabrib@math.uio.no}

% \cgsn{Research Council of Norway: Center for Biomedical Computing at Simula Research Laboratory
%       and Center for Cardiological Innovation at Oslo University Hospital}{179578 and 203489}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
% \keywords{cardiac mechanics; data assimilation; patient-specific; adjoint; dyssynchrony}
\section*{High Resolution Data Assimilation of Cardiac Mechanics}

\begin{center}
G. Balaban\footnotemark,
H. Finsberg\footnotemark[\value{footnote}], 
H. Odland, M. E. Rognes,
S. Ross, J. Sundnes, and
S. Wall
\end{center}

% \begin{center}
% G. Balaban\textsuperscript{\footnotemark 1,2,5},
% H. Finsberg\textsuperscript{\footnotemark[\value{footnote}], 1,2,5}, 
% H. Odland\textsuperscript{4,7}, M. E. Rognes\textsuperscript{1,3},
% S. Ross\textsuperscript{4,5}, J. Sundnes\textsuperscript{1,2,5},
% S. Wall\textsuperscript{1,5, 6} 
% \end{center}

% \begin{enumerate}
%   \itemsep-0.5em 
% \item Simula Research Laboratory, P.O. Box 134 1325 Lysaker, Norway
% \item Department of Informatics, University of Oslo, P.O. Box 1080, Blindern 0316 Oslo, Norway
% \item Department of Mathematics, University of Oslo, P.O Box 1053, Blindern 0316 Oslo, Norway
% \item Faculty of Medicine, University of Oslo, P.O. Box 1078 Blindern, 0316 Oslo, Norway 
% \item Center for Cardiological Innovation, Songsvannsveien 9, 0372 Oslo, Norway
% \item Department of Mathematical Science and Technology, Norwegian University of Life Sciences,
%   Universitetstunet 3 1430 Ås, Norway
% \item Department of Pediatrics, Oslo University Hospital, PO Nydalen, Oslo, Norway
% \end{enumerate}

@@

\appendix

\section{Sensitivity of elastic $a$ parameter to error in atrial systolic volume measurements}
In order to test the sensitivity of our estimated $a$ parameter to
uncertainty in volume measurements we have carried out a series of estimations with
various levels of volume perturbation. 
We generated clean volume data using the computational model using $a = 0.435$ kPa, 
the optimal value obtained from the clinical data. Perturbations in volume increases
of sizes $\pm 5,15, 25 \%$ were added to this data,
which were then used as target for optimization. The resulting $a$ values and perturbations
are shown in Table~\ref{tab:passive_synth_opt}. The largest perturbations resulted
in the $a$ values 0.494 kPa and 0.384 kPa, representing circa $\pm \%13$ change from the original $a$ value.

@@

\section{Sensitivity of estimated parameters to spring constant $k$.}
\label{sec:k_sense}
The spring boundary condition that we used at the ventricular base has a significant
effect on the simulated cavity volumes calculated by the model. This is due to the 
cross-sectional area of the cavity being large at the ventricular base. Therefore 
we can expect the choice of $k$ to have an effect on the optimal parameters calculated 
by our data assimilation.

@@

\section{Effect of mesh resolution on estimated contraction $\gamma$ at P1 resolution}
\label{sec:mesh_res}
Ventricular meshes were generated by Gmsh
\cite{geuzaine2009gmsh} with three different resolutions controlled by the parameter
'Mesh.CharacteristicLengthFactor'.  This parameter
was given the values $1.0$, $0.65$ and $0.45$ which resulted in meshes with
$549$, $1262$ and $2261$ vertices respectively. Using the three meshes
we estimated contraction fields from the in-vivo data. The average value of
$\gamma$ is shown for these three cases in Figure~\ref{fig:mesh_conv_gamma}. 
Fit quality is compared in Table~\ref{tab:mesh_conv_opt_misfit}.

@@

\section{Sensitivity of Contraction Size to choices of $\alpha$ and $\lambda$}
\label{sec:sense_alpha_lambda}
Based on the trade-off curves in Figure~\ref{fig:lcurves} we chose 
the optimization weights $\alpha = 0.95$ and $\lambda = 0.01$ for the personalization
of our wall motion model to the in-vivo data. In order to show the effect of these 
choices on the optimized contraction field $\gamma$ we have varied
the $\alpha$ and $\lambda$ values and plotted the spatial averages of the resulting
contraction fields. The results show that the amount of contraction tends to increase
proportionally to both $\alpha$ and $\lambda$ beyond the thresholds $\alpha = 0.5$ and
$\lambda = 0.001$.

@@

\section{Estimation of noise in echo speckle tracking strain measurements}
\label{sec:strain_noise_est}
In order to increase the relevance of the synthetic tests we
considered a set of regional strains that contained noise. This
noise was modelled as an additive Gaussian process in order to imitate
the accumulation of tracking errors in EchoPac's image based strain
calculations. The mean and variance of a summand in the Gaussian
process were estimated from our in-vivo strain data of a single patient. From this data 
the sample means and variances of
the drift values were divided by the number of measurement points in order to approximate the noise in 
a single measurement. The mean and covariance of this single measurement point noise are given in
Table~\ref{tab:noise_covar_mean}.Theoretically, error free strain
curves would have no drift given stable conditions in the heart.
This motivates the use of the drift values in order to approximate 
the tracking error.

>>>>

\graphicspath{{chapters/paper1/figures/}}

@@

@@

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\section*{High Resolution Data Assimilation of Cardiac Mechanics}

% \begin{center}
G. Balaban\footnotemark,
H. Finsberg\footnotemark[\value{footnote}], 
H. Odland, M. E. Rognes, \\
S. Ross, J. Sundnes, and
S. Wall
% \end{center}

@@

\section{Appendix}

\subsection{Sensitivity of elastic $a$ parameter to error in atrial systolic volume measurements}
In order to test the sensitivity of our estimated $a$ parameter to
uncertainty in volume measurements we have carried out a series of estimations with
various levels of volume perturbation. 
We generated clean volume data using the computational model using $a = 0.435$ kPa, 
the optimal value obtained from the clinical data. Perturbations in volume increases
of sizes $\pm 5,15, 25 \%$ were added to this data,
which were then used as target for optimization. The resulting $a$ values and perturbations
are shown in Table~\ref{tab:passive_synth_opt}. The largest perturbations resulted
in the $a$ values 0.494 kPa and 0.384 kPa, representing circa $\pm \%13$ change from the original $a$ value.

@@

\subsection{Sensitivity of estimated parameters to spring constant $k$.}
\label{sec:k_sense}
The spring boundary condition that we used at the ventricular base has a significant
effect on the simulated cavity volumes calculated by the model. This is due to the 
cross-sectional area of the cavity being large at the ventricular base. Therefore 
we can expect the choice of $k$ to have an effect on the optimal parameters calculated 
by our data assimilation.

@@

\subsection{Effect of mesh resolution on estimated contraction $\gamma$ at P1 resolution}
\label{sec:mesh_res}
Ventricular meshes were generated by Gmsh
\cite{geuzaine2009gmsh} with three different resolutions controlled by the parameter
'Mesh.CharacteristicLengthFactor'.  This parameter
was given the values $1.0$, $0.65$ and $0.45$ which resulted in meshes with
$549$, $1262$ and $2261$ vertices respectively. Using the three meshes
we estimated contraction fields from the in-vivo data. The average value of
$\gamma$ is shown for these three cases in Figure~\ref{fig:mesh_conv_gamma}. 
Fit quality is compared in Table~\ref{tab:mesh_conv_opt_misfit}.

@@

\subsection{Sensitivity of Contraction Size to choices of $\alpha$ and $\lambda$}
\label{sec:sense_alpha_lambda}
Based on the trade-off curves in Figure~\ref{fig:lcurves} we chose 
the optimization weights $\alpha = 0.95$ and $\lambda = 0.01$ for the personalization
of our wall motion model to the in-vivo data. In order to show the effect of these 
choices on the optimized contraction field $\gamma$ we have varied
the $\alpha$ and $\lambda$ values and plotted the spatial averages of the resulting
contraction fields. The results show that the amount of contraction tends to increase
proportionally to both $\alpha$ and $\lambda$ beyond the thresholds $\alpha = 0.5$ and
$\lambda = 0.001$.

@@

\subsection{Estimation of noise in echo speckle tracking strain measurements}
\label{sec:strain_noise_est}
In order to increase the relevance of the synthetic tests we
considered a set of regional strains that contained noise. This
noise was modelled as an additive Gaussian process in order to imitate
the accumulation of tracking errors in EchoPac's image based strain
calculations. The mean and variance of a summand in the Gaussian
process were estimated from our in-vivo strain data of a single patient. From this data 
the sample means and variances of
the drift values were divided by the number of measurement points in order to approximate the noise in 
a single measurement. The mean and covariance of this single measurement point noise are given in
Table~\ref{tab:noise_covar_mean}.Theoretically, error free strain
curves would have no drift given stable conditions in the heart.
This motivates the use of the drift values in order to approximate 
the tracking error.
